---
title: "Happy Moments"
author: Lujia Wang (lw2772)
output:
  html_document: default
---

```{r, warning=FALSE, message=FALSE,echo=FALSE}
packages.used=c("tm", "tidyverse", "tidytext", 
                "DT", "scales", "wordcloud2",
                "gridExtra", "ngram")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}
library(tm)
library(tidyverse)
library(tidytext)
library(DT)
library(scales)
library(wordcloud2)
library(gridExtra)
library(ngram)
```



```{r load data, warning=FALSE, message=FALSE,echo=FALSE}
# Load demographic data, processed text data and senselabel data
hm_data <- read_csv("../output/processed_moments.csv")
demo.urlfile<-'https://raw.githubusercontent.com/rit-public/HappyDB/master/happydb/data/demographic.csv'
sense.urlfile <- 'https://raw.githubusercontent.com/rit-public/HappyDB/master/happydb/data/senselabel.csv'
sense_data <- read_csv(sense.urlfile,progress = FALSE)
demo_data <- read_csv(demo.urlfile)
demo_data$age <- as.numeric(demo_data$age) # change age datatype into numeric
```


```{r combining data, warning=FALSE, message=FALSE,echo=FALSE}
# Combine both the data sets and keep the required columns for analysis
# We select a subset of the data that satisfies specific row conditions.
hm_data <- hm_data %>%
  inner_join(demo_data, by = "wid") %>%
  select(id,
         hmid,
         wid,
         cleaned_hm,
         gender, 
         marital, 
         parenthood,
         reflection_period,
         num_sentence,
         age, 
         country, 
         ground_truth_category,
         predicted_category,
         text) %>%
  mutate(count = sapply(hm_data$text, wordcount)) %>%
  filter(gender %in% c("m", "f")) %>%
  filter(marital %in% c("single", "married")) %>%
  filter(age < 120 & age>5) %>%  #some ages are unreasonable
  filter(parenthood %in% c("n", "y")) %>%
  filter(reflection_period %in% c("24h", "3m")) %>%
  #filter(country=="USA") %>%              #some languages from other countries are weird
  mutate(reflection_period = fct_recode(reflection_period, 
                                        months_3 = "3m", hours_24 = "24h"))

```

### How many sentences do you use to describe a happy moment?

Most happy moments is only in one sentence. Simple but powerful, isn't it?

However, A few happy memory is a long story and we need more words to express our feelings. Young people, especially male around age of 25, like to write extra long paragraphs describing their happy moments.

*There are some outliers in the data set. Someone copied a story and had multiple same entries. I have removed these answers from the data set.*

```{r,warning=FALSE, message=FALSE,echo=FALSE}
# remove unrelated answers
hm_data <- hm_data%>%filter(wid!=455&wid!=508)
# calculated number of sentences distribution by gender
num_sentence_dist <- hm_data %>% group_by(num_sentence,gender) %>% summarise(n=length(id))
num_sentence_dist$num_sentence[num_sentence_dist$num_sentence > 5] <- '>5'
num_sentence_dist$num_sentence <- factor(num_sentence_dist$num_sentence,levels = c(">5", "5", "4","3","2","1"))

# Age and number of sentence scatter plot
p1 <- ggplot(hm_data, aes(x=age, y=num_sentence,color=gender)) +
  geom_point(size=1,alpha=0.7) +
  labs(y='number of sentences',title="Number of sentences vs Age\n") +
  theme_bw() +
  theme(plot.title=element_text(size=10, face="bold"))
# number of sentence distribution by gender stacked bar plot
p2 <- ggplot(num_sentence_dist, aes(x=gender, y=n, fill=num_sentence))+
  geom_bar(width = 0.7, stat = "identity")+
  scale_fill_brewer(palette="GnBu") +
  labs(fill='number of \nsentences',
       title="Distribution of number \nof sentences") +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title=element_text(size=10, face="bold"))

grid.arrange(p1, p2, nrow = 2, ncol = 3, layout_matrix=rbind(c(1,1,2),c(1,1,2)))
```

***

# Happiness in ONE line!

In order to avoid randomness, I focus on those moments written in one line.

### What are they about?

Achievements and affection are two major sources of our happiness. A possible guess is that achievements are from our jobs and work, and feelings of affection are from our friends and family.

```{r, warning=FALSE,message=FALSE,echo=FALSE}
#filter happy moments in only one sentence.
one.line <- hm_data %>% filter(num_sentence==1)
# plot a circle bar plot to show distributions of happy moments by category. 
ggplot(one.line, aes(x=predicted_category,fill=predicted_category)) + 
  geom_bar() +
  ylim(-20000,28000) +
  coord_polar(start = 0)+
  theme_minimal() +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        plot.title=element_text(size=15, face="bold")) +
  labs(title='One-line happy moments by category',fill="category")
```

A lot of happy time happen between friends, family, wife, husband, daughter, son and ourselves.

```{r,warning=FALSE,message=FALSE,echo=FALSE}
bagOfWords <-  one.line %>% unnest_tokens(word, text)
word_count <- bagOfWords%>% count(word, sort = TRUE)
word_count %>% top_n(20, n) %>%
ggplot(aes(x=reorder(word,n),y=n))+
  geom_bar(stat = "identity",fill="skyblue")+
  theme_minimal() + coord_flip() +
  labs(title="Most frequent words in HappyDB",
       y="frequency",x="words") +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank(),
        plot.title=element_text(size=15, face="bold"))
#wordcloud2(word_count,ellipticity = 0.1,size=0.6,color = "random-light")
```

Now, I want oanalyze frequencies of words at different positions in a sentence.

  + Happiness begins with...?

Obviously, most happy moments focus on ourselves and things happen on us.

  + Who makes us happy?

Naturally, we can always find happiness with our family and friends. I also notice an interesting fact is that son/daughter appears more often than mother/father/mom/dad. Animals, including human, care about their children much more than their parents. This rule is reflected here.

  + What things makes us happy?

Some happiness are from work like a job promotion, a salary raise. Some happiness are just small things like a tasty dinner,a good movie, or a book. Some happiness are from our pets. Dog is the most beloved pet for Americans.

```{r Happiness begins with,warning=FALSE,message=FALSE,echo=FALSE}
sense <- one.line %>% select("hmid","predicted_category") %>% inner_join(sense_data, by="hmid")
sense$supersenseLabel[sense$lowercaseLemma=="family"] <- "n.person"
sense$POS[sense$lowercaseLemma=="i"] = "PRON"
sense$supersenseLabel[sense$lowercaseLemma=="time"] <- "n.time"
sense$supersenseLabel[sense$lowercaseLemma=="year"] <- "n.time"

first_word_count <- sense %>%filter(tokenOffset==1) %>% count(lowercaseLemma, sort = TRUE)
p1 <- ggplot(first_word_count[1:20,], aes(x=reorder(lowercaseLemma,n),y=n))+
  geom_bar(stat = "identity",fill="Orange")+
  theme_minimal() + coord_flip() +
  labs(title="Happiness begins with...?",
       y="frequency",x="first words in sentences") +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank(),
        plot.title=element_text(size=10, face="bold"))
  
```

```{r Who makes us happy,warning=FALSE,message=FALSE,echo=FALSE}
person_word_count <- sense %>%
  filter(POS=="NOUN"&supersenseLabel=="n.person")%>%
  count(lowercaseLemma, sort = TRUE)
p2 <- ggplot(person_word_count[1:20,], 
       aes(x=reorder(lowercaseLemma,n),y=n))+
  geom_bar(stat = "identity", fill="chocolate")+
  theme_minimal() + coord_flip() +
  labs(title="Who makes us happy?",y="frequency",x="words of people") +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank(),
        plot.title=element_text(size=10, face="bold"))
  
```

```{r What thing makes us happy,warning=FALSE,message=FALSE,echo=FALSE}
noun_word_count <- sense %>%
  filter(POS=="NOUN"&supersenseLabel!="n.person"&supersenseLabel!="n.time")%>%
  count(lowercaseLemma, sort = TRUE)
p3 <- ggplot(noun_word_count[1:20,], 
       aes(x=reorder(lowercaseLemma,n),y=n))+
  geom_bar(stat = "identity",fill="skyblue")+
  theme_minimal() + coord_flip() +
  labs(title="What things make us happy?",
       y="frequency",x="word of things") +
  theme(axis.text.x = element_blank(),
        panel.grid=element_blank(),
        plot.title=element_text(size=10, face = "bold"))
  
grid.arrange(p1, p2, p3, ncol=3)
```

***

## Sentiment Analysis

##### Sentiment score of different types of people

I calculated the average sentiment score by their gender, marital and parenthood status. Surprisingly, single dads have the highest postivity score. Another interesting fact is that for female, married women have similar or slight higher positivity score with unmarried women. However, the scenerio is opposite for male. Perhaps it is because women enjoy marriage more than men.

Having a child also infuences the positivity score of male and female in different ways. A married women who has kids has a lower score than a married women who does't have kids. Men who have kids have a positivity score higher than those who don't have kids.

```{r,message=FALSE,warning=FALSE,echo=FALSE}
words_by_demo <- bagOfWords %>%
  count(gender,marital,parenthood, word, sort = TRUE) %>%
  ungroup()

sentiments_by_demo <- words_by_demo %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(gender,marital,parenthood) %>% summarise(score = sum(score * n) / sum(n))

ggplot(sentiments_by_demo, aes(x=marital,y=score,fill=parenthood)) +
  geom_col(position = "dodge") +
  facet_wrap(~gender) +
  theme_bw() +
  scale_fill_brewer(palette="Blues") +
  theme(plot.title=element_text(size=15, face = "bold")) +
  labs(title="Sentiment Score by Gender")

```


##### Sentiment score of each category

I also calculate the average sentiment score of different categories. Though not so much entries of happy moments are about nature, nature has the highest positivity score among all categories. Because there are more **nice** words when describing the nature, such as beautiful, warm and so on. We will see that in a moment.


```{r,message=FALSE,warning=FALSE,echo=FALSE}
words_by_category <- bagOfWords %>%
  count(predicted_category, word, sort = TRUE) %>%
  ungroup()

sentiments_by_category <- words_by_category %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(predicted_category) %>%
  summarize(score = sum(score * n) / sum(n))

sentiments_by_category %>%
  mutate(category = reorder(predicted_category, score)) %>%
  ggplot(aes(category, score)) +
  geom_col(show.legend = FALSE,fill="lightblue") +
  coord_flip() +
  theme_bw() +
  theme(plot.title=element_text(size=15, face = "bold")) +
  labs(y="Average sentiment score",
       title="Average Sentiment Score by Category")
```

##### What words have the most effect on sentiment scores overall?

Here I analyze the contribution (sentiment score weighted by occurrences) of each words and select the top 20 words.

```{r,message=FALSE,warning=FALSE,echo=FALSE}
sentiments_by_word <- bagOfWords %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(word) %>%
  summarize(occurences = n(),
            contribution = sum(score))

sentiments_by_word %>%
  top_n(20, abs(contribution)) %>%
  mutate(word = reorder(word, contribution)) %>%
  ggplot(aes(word, contribution, fill = contribution > 0)) +
  scale_fill_brewer() +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  theme_bw() +
  theme(plot.title=element_text(size=15, face = "bold"))+
  labs(title="Words with most effect on sentiment score")
```

##### What word havs the most effect on sentiment scores within each category

After analyzing the overall contribution of each word, now I evaluate their contribution within each category. **love** appears in almost every category, and it makes sense.

```{r,message=FALSE,warning=FALSE,echo=FALSE,fig.height=10}
top_sentiment_words <- words_by_category %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  mutate(contribution = score * n / sum(n))
  
top_sentiment_words %>%
  group_by(predicted_category) %>%
  top_n(5, abs(contribution)) %>%
  ungroup() %>%
  mutate(word=reorder(paste(word,predicted_category,sep = "__"), contribution)) %>%
  ggplot(aes(word, contribution, fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  facet_wrap(~ predicted_category, scales = "free", nrow = 4) +
  coord_flip() +
  theme_bw() +
  theme(plot.title=element_text(size=15, face = "bold")) +
  labs(title = "Top five words with highest contribution in each category")
  
```

***

## Bigram Analysis

### What is my favorite?

Here I count top ten words preceded by word *favorite/favourite*. Interestingly, SEVEN of them, including the top ones, are about eating. Those words are *restaurant*, *food*, *meal*, *dish*, *snack*, *sweet*, *dinner*. Apparently, Eating is so happy! 

```{r,echo=FALSE}
#bigram analysis
bigrams <- one.line %>% filter(count!=1) %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)
#count bigram frequency
bigram_counts <- bigrams %>%
  count(predicted_category, bigram, sort = TRUE) %>%
  ungroup() %>%
  separate(bigram, c("word1", "word2"), sep = " ")

# favorite words
first_word <- c("favorite","favourite")
# top 10 words followed favorite
bigram_counts %>%
  filter(word1 %in% first_word) %>%
  count(word1, word2, wt = n, sort = TRUE) %>%
  group_by(word1) %>%
  top_n(10, nn) %>%
  ungroup() %>%
  mutate(word2 = reorder(paste(word2, word1, sep = "__"), nn)) %>%
  ggplot(aes(word2, nn)) +
  geom_col(show.legend = FALSE,fill="orange") +
  facet_wrap(~ word1, scales = "free", nrow = 1) +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(x="Words preceded by favorite",
       y="Number of occurrences",
       title="My favorite is ...") +
  theme_bw() +
  theme(plot.title=element_text(size=15, face = "bold")) +
  coord_flip()
```



### Dinner, my favorite eating time!

Of all three meals, we love dinners, expecially with our beloved friends and family. If you are struggling or frustrated about something, having a decent dinner may be the quickest way to get you happy.

```{r,message=FALSE,warning=FALSE,echo=FALSE,fig.height=2}
food.dict <- read_csv('https://raw.githubusercontent.com/rit-public/HappyDB/master/happydb/data/topic_dict/food-dict.csv', col_names = FALSE)
food <- as.vector(food.dict$X1)
food.df <- bagOfWords %>% 
  filter(word %in% food) %>% select(id) %>% unique() %>%
  left_join(one.line, by=c("id"="id"))
bagOfWords_food <- food.df %>% unnest_tokens(word, text)
word_count_food <- bagOfWords_food %>% count(word, sort = TRUE)
wordcloud2(word_count_food,ellipticity = 0.1,size=0.6)
```



### Yummy! I am so happy when I am eating/drinking ...

Here are most frequent words about food and drinks in the set. **Coffee** and **Pizza** are the best food/drink to cheer one up. Sweet dessert like ice cream, chocolate, cake and candy can also give a sense of sweetness to our lives. 

Next time, if you are unhappy, maybe you are just hungry. Order a pizza and invite your friend to share it with you, everything will be okay :) 

I also noted that among these food, the only one that is more authentic is Sushi. Maybe that's reason why sushi is so polular worldwide.


```{r,echo=FALSE,message=FALSE,warning=FALSE}
# join happy moments on food topic with the sense data
sense.food <- food.df %>% select("hmid") %>% left_join(sense,by="hmid")
food_word_count <- sense %>%
  filter(supersenseLabel=="n.food"&
           (!lowercaseLemma%in%c("food","dinner","lunch","breakfast","drink")))%>%
  count(lowercaseLemma, sort = TRUE)
# Top 20 words about food
ggplot(food_word_count[1:20,], 
       aes(x=reorder(lowercaseLemma,n),y=n))+
  geom_bar(stat = "identity",fill="Orange")+
  theme_minimal() + coord_flip() +
  theme(axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.grid = element_blank(),
        plot.title=element_text(size=15, face = "bold")) +
  labs(title="What food or drinks makes you happy?",
       x="food/drinks",y="frequency")
```

## Future Analysis

Some future anaysis I can think of right now is 

+ topic modeling, categorized happy moments into topics like exercise, shopping, job, eating, pets, video games...

+ Differences between happy memories reflected from short term and long term.

+ Based on different segmentatio of people, analyze what activities they like to do.

